---
title: "Weight Lifting Exercise Dataset"
output: html_document
---

Here a machine learning algorithm is developed with a training data consisting of 19622 entries and 160 variable.The purpose is to develop a model that identifies five different training styles. This models should be ultimately used to predict training styles of 20 test casses.

The very first process is that the data is read in R

```{r, reading data}
train = read.csv("C:\\Users\\Deepak\\Documents\\Practical Machine Learning\\pml-training.csv")
test = read.csv("C:\\Users\\Deepak\\Documents\\Practical Machine Learning\\pml-testing.csv")
```

Then they are cleaned. This is done in two steps. 
First the variable with all NA were deleted, from both training and test data sets. I changed the names to train1 and test1. Then other variables  x,user_name, raw_timestamp_part_1, cvtd_timestamp, new_window variable are taken out becasue, they doen't help in prediction. 

```{r, cleaning data}
myvars <- colSums(is.na(test))<nrow(test)
train1<- train[,myvars]
test1<- test[,myvars]
train1<- train1[c(-1,-2,-3,-4,-5,-6)]
test1<- test1[c(-1,-2,-3,-4,-5,-6)]
```

Then prediction process is started. I used the random forest method to develop the model using the tainnning set (train1). However, because my computer could not take big data, I could run this model by taking maximum of 2000 observations.Even with this it took about 20 minutes to run my modFid command given below.


```{r, developing the model}
library(caret)
set.seed(9850)
g<- runif(nrow(train1))
train1<- train1[order(g),]

modFit <- train(classe~.,data=train1[1:2000,], method="rf", number=5, allowParallel=T)
print(modFit$finalModel)

```

Here overall within the sample error was 3.8%. I used this to predict the test data as following.

```{r, predicting with the model}
pr2<-predict(modFit,newdata=test1)
pr2

```


My prediction was 95% accurate- meaning out of 20, 19 predictions turned out to be correct. The error could be slightly higher than the insample but it is not significant. 